---
title: "Surdek Milestone Three"
author: "msurdek"
date: "9/22/2022"
output: html_document
---
## markdown setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(width = 100)
```

## load required packages

```{r, message=FALSE, warnings=FALSE, results=FALSE, include = FALSE}
library(dplyr)
library(ggplot2)
library(tidymodels)
library(recipes)
library(forcats)
library(tidytext)
library(readr)
library(pastecs)
library(moments)
library(corrplot)
library(car)
library(janitor)
library(caret)
library(glmnet)
library(themis)
library(usmap)
```

## FUNCTION get csa info

```{r, include = FALSE}
get_csa_info <- function(df) {
  
  df |> 
    
    mutate(CSA_main = substr(CSA,1,3)) |> 
    
    mutate(CSA_state = case_when(
      CSA_main == 'NYC' ~ 'NY', 
      CSA_main == 'LAX' ~ 'CA',
      CSA_main == 'DAL' ~ 'TX', 
      CSA_main == 'MIA' ~ 'FL', 
      CSA_main == 'APC' ~ 'MD',
      CSA_main == 'SFR' ~ 'CA', 
      CSA_main == 'NEV' ~ 'NV',
      CSA_main == 'SAN' ~ 'TX',
      CSA_main == 'BOS' ~ 'MA',
      CSA_main == 'DET' ~ 'MI',
      CSA_main == 'CHI' ~ 'IL',
      CSA_main == 'OHI' ~ 'OH',
      CSA_main == 'FLN' ~ 'FL',
      CSA_main == 'HOU' ~ 'TX',
      CSA_main == 'PHI' ~ 'PA',
      CSA_main == 'SEA' ~ 'WA',
      CSA_main == 'ATL' ~ 'GA',
      CSA_main == 'KCY' ~ 'MO',
      CSA_main == 'NCR' ~ 'VA',
      CSA_main == 'AIR' ~ 'NC',
      CSA_main == 'HAR' ~ 'CT',
      CSA_main == 'NNY' ~ 'NY',
      CSA_main == 'STL' ~ 'MO',
      CSA_main == 'DEN' ~ 'CO',
      CSA_main == 'MIL' ~ 'WI',
      CSA_main == 'MIN' ~ 'MN',
      CSA_main == 'OMA' ~ 'NE',
      CSA_main == 'LOU' ~ 'KY',
      CSA_main == 'OKC' ~ 'OK',
      CSA_main == 'IND' ~ 'IN',
      CSA_main == 'NSH' ~ 'TN',
      CSA_main == 'PHX' ~ 'AZ',
      CSA_main == 'PIT' ~ 'PA',
      CSA_main == 'HWI' ~ 'HI',
      CSA_main == 'NMX' ~ 'NM',
      CSA_main == 'NOL' ~ 'LA',
      CSA_main == 'OHH' ~ 'OH',
      CSA_main == 'AWI' ~ 'WI',
      CSA_main == 'BIR' ~ 'AL',
      CSA_main == 'LAU' ~ 'MS',
      CSA_main == 'INH' ~ 'IN',
      CSA_main == 'IPM' ~ 'MI',
      CSA_main == 'SEW' ~ 'WA',
      CSA_main == 'SHE' ~ 'MD',
      CSA_main == 'SLC' ~ 'UT')) |> 
    
    mutate(CSA_region = case_when(
      CSA_state == 'AL' ~ 'Southeast', 
      CSA_state == 'AZ' ~ 'Southwest',
      CSA_state == 'CA' ~ 'West', 
      CSA_state == 'CO' ~ 'West', 
      CSA_state == 'CT' ~ 'Northeast', 
      CSA_state == 'FL' ~ 'Southeast',
      CSA_state == 'GA' ~ 'Southeast', 
      CSA_state == 'HI' ~ 'West', 
      CSA_state == 'IL' ~ 'Midwest',
      CSA_state == 'IN' ~ 'Midwest',
      CSA_state == 'KY' ~ 'Midwest', 
      CSA_state == 'LA' ~ 'South', 
      CSA_state == 'MA' ~ 'Northeast',
      CSA_state == 'MD' ~ 'East', 
      CSA_state == 'MI' ~ 'North',
      CSA_state == 'MN' ~ 'North', 
      CSA_state == 'MO' ~ 'Midwest',
      CSA_state == 'MS' ~ 'South', 
      CSA_state == 'NC' ~ 'East', 
      CSA_state == 'NE' ~ 'Midwest',
      CSA_state == 'NM' ~ 'Southwest', 
      CSA_state == 'NV' ~ 'West', 
      CSA_state == 'NY' ~ 'Northeast',
      CSA_state == 'OH' ~ 'Midwest', 
      CSA_state == 'OK' ~ 'South',
      CSA_state == 'PA' ~ 'East',
      CSA_state == 'TN' ~ 'Midwest',
      CSA_state == 'TX' ~ 'South',
      CSA_state == 'UT' ~ 'West',
      CSA_state == 'VA' ~ 'East',
      CSA_state == 'WA' ~ 'Northwest',
      CSA_state == 'WI' ~ 'North')) |> 
    
    mutate(CSA_region_SEW = case_when(
      CSA_region == 'Southeast' ~ 1, 
      CSA_region == 'East' ~ 1,
      CSA_region == 'West' ~ 1, 
      CSA_region == 'Southwest' ~ 1, 
      TRUE ~ 0)) |> 
    
    mutate(CSA_region_SW = case_when(
      CSA_region == 'Southwest' ~ 1, 
      TRUE ~ 0)) |> 
    
    mutate(CSA_region_NS = case_when(
      CSA_region == 'North' ~ 1,
      CSA_region == 'South' ~ 1,
      TRUE ~ 0))
  
}
```

## FUNCTION get derived attributes

```{r, include = FALSE}
get_attributes <- function(df) {
  
  df |> 
    mutate(CHANGEM_mod = case_when(CHANGEM >= 100 & CHANGEM <= 300 ~ 1, TRUE ~ 0)) |> 
    mutate(INCOME_medhigh = as.factor(case_when(INCOME >= 5 & INCOME <= 7 ~ 1, TRUE ~ 0))) |> 
    mutate(CREDITDEGY = case_when(CREDIT_LABEL == 'DE' | CREDIT_LABEL == 'GY' ~ 1, TRUE ~ 0)) |> 
    mutate(REFERRAL = as.factor(case_when(REFER >= 1 ~ 1, TRUE ~ 0))) |> 
    mutate(CREDIT_ADJUSTMENT = as.factor(case_when(CREDITAD >= 1 ~ 1, TRUE ~ 0))) |> 
    mutate(MOU_low = case_when(MOU <= 220 ~ 1, TRUE ~ 0)) |> 
    mutate(OCC_NPS = case_when(OCC_LABEL == 'NONE' | OCC_LABEL == 'STUDENT' |OCC_LABEL == 'PROFESSIONAL' ~ 1, OCC_LABEL == 'CLERICAL' | OCC_LABEL == 'CRAFT' |OCC_LABEL == 'HOMEMAKER' | OCC_LABEL == 'RETIRED' | OCC_LABEL == 'SELF-EMPLOYED' ~ 0))
  
}
```

## FUNCTION zeros to nas

```{r, include = FALSE}
zeros_to_nas <- function(df) {
  
  df <- df |> 
    mutate_at(c('AGE1'), ~na_if(., 0)) |> 
    mutate_at(c('AGE2'), ~na_if(., 0)) |> 
    mutate_at(c('INCOME'), ~na_if(., 0)) |> 
    mutate_at(c('SETPRC'), ~na_if(., 0))
  
  return(df)
  
}
```

## FUNCTION read churn file

```{r, include = FALSE}
read_churn_file <- function(path) {
  
  data <- read_csv(path) |> 
    rename(PRZM_NUM = 'Column 45') |> 
    mutate(CHURN_TARGET = as.factor(CHURN)) |> 
    mutate(CHURN_INT = CHURN) |> 
    mutate(MARRY_LABEL = case_when (MARRYYES == 1 ~ 'YES', MARRYNO == 1 ~ 'NO',MARRYUN == 1 ~ 'UNKNOWN')) |> 
    mutate(PRIZM_LABEL = case_when(PRIZMRUR == 1 ~ 'RURAL', PRIZMUB == 1 ~ 'URBAN', PRIZMTWN == 1 | PRZM_NUM == 3 ~ 'TOWN', TRUE ~ 'UNKNOWN')) |> 
    mutate(CREDIT_LABEL = case_when(CREDITA == 1 ~ 'A', CREDITAA == 1 ~ 'AA', CREDITB == 1 ~ 'B', CREDITC == 1 ~ 'C', CREDITDE == 1 ~ 'DE', CREDITGY == 1 ~ 'GY', CREDITZ == 1 ~ 'Z')) |> 
    get_csa_info() |> 
    select(!c(CALIBRAT, CHURNDEP, CHURN))
  
}
```

## FUNCTION impute missing values

```{r, include = FALSE}
impute_values <- function(df) {
  
  df |> 
    mutate(REVENUE = replace_na(REVENUE,revenue_median)) |> 
    mutate(MOU = replace_na(MOU,minutes_median)) |> 
    mutate(RECCHRGE = replace_na(RECCHRGE,recchrge_median)) |> 
    mutate(DIRECTAS = replace_na(DIRECTAS,directas_median)) |> 
    mutate(OVERAGE = replace_na(OVERAGE,overage_median)) |> 
    mutate(ROAM = replace_na(ROAM,roaming_median)) |> 
    mutate(CHANGEM = replace_na(CHANGEM,changem_zero)) |> 
    mutate(CHANGER = replace_na(CHANGER,changer_zero)) |> 
    mutate(AGE1 = replace_na(AGE1,age1_median)) |> 
    mutate(AGE2 = replace_na(AGE2,age2_median)) |> 
    mutate(INCOME = replace_na(INCOME,income_median)) |> 
    mutate(SETPRC = replace_na(SETPRC,setprice_median))
  
}
```

## read data files

```{r}
data_churn <- read_churn_file("/home/msurdek/Documents/School/SNHU/DAT 690 Capstone in Data Analytics/Assignments/Project/Data/Churn Data.csv")

data_churn_test <- read_churn_file("/home/msurdek/Documents/School/SNHU/DAT 690 Capstone in Data Analytics/Assignments/Project/Data/Churn Verification Data.csv")
```

## calculate values to impute

```{r, include = FALSE}
revenue_median <- median(data_churn$REVENUE |> na.omit())
minutes_median <- median(data_churn$MOU |> na.omit())
recchrge_median <- median(data_churn$RECCHRGE |> na.omit())
directas_median <- median(data_churn$DIRECTAS |> na.omit())
overage_median <- median(data_churn$OVERAGE |> na.omit())
roaming_median <- median(data_churn$ROAM |> na.omit())
changem_zero <- 0
changer_zero <- 0
age1_median <- median(data_churn$AGE1 |> na.omit())
age2_median <- median(data_churn$AGE2 |> na.omit())
income_median <- median(data_churn$INCOME |> na.omit())
setprice_median <- median(data_churn$SETPRC |> na.omit())
```

## impute missing values

```{r}
data_churn <- data_churn |> impute_values()

data_churn_test <- data_churn_test |> impute_values()
```

## add derived attributes

```{r}
data_churn <- data_churn |> get_attributes()

data_churn_test <- data_churn_test |> get_attributes()
```

## prep principal components

```{r, include = FALSE}
# USAGE

usage_pca_vars <- c('DIRECTAS','OVERAGE','ROAM','CHANGEM','DROPVCE','BLCKVCE','UNANSVCE','THREEWAY','MOUREC','OUTCALLS','INCALLS','PEAKVCE','OPEAKVCE','DROPBLK','CALLFWDV','CALLWAIT','DROPVCE','DROPBLK','BLCKVCE')

usage_pca_vars <- c('DIRECTAS','OVERAGE','ROAM','CALLFWDV','REVENUE','RECCHRGE')

usage_pca_data <- data_churn |>
  select(c('CUSTOMER','CHURN_TARGET',all_of(usage_pca_vars)))

usage_pca_rec <- recipe(~., data = usage_pca_data) %>%
  update_role(CUSTOMER, CHURN_TARGET, new_role = "id") %>%
  step_normalize(all_predictors()) %>%
  step_pca(all_predictors())

usage_pca_prep <- prep(usage_pca_rec)

usage_pca_values <- juice(usage_pca_prep) |> 
  rename(USAGEPC1 = "PC1", USAGEPC2 = "PC2", USAGEPC3 = "PC3", USAGEPC4 = "PC4", USAGEPC5 = "PC5")

usage_pca_values_test <- bake(usage_pca_prep, data_churn_test) |> 
  rename(USAGEPC1 = "PC1", USAGEPC2 = "PC2", USAGEPC3 = "PC3", USAGEPC4 = "PC4", USAGEPC5 = "PC5")

# COVERAGE

coverage_pca_vars <- c('DROPVCE','DROPBLK','BLCKVCE')

coverage_pca_data <- data_churn |>
  select(c('CUSTOMER','CHURN_TARGET',all_of(coverage_pca_vars)))

coverage_pca_rec <- recipe(~., data = coverage_pca_data) %>%
  update_role(CUSTOMER, CHURN_TARGET, new_role = "id") %>%
  step_normalize(all_predictors()) %>%
  step_pca(all_predictors())

coverage_pca_prep <- prep(coverage_pca_rec)

coverage_pca_values <- juice(coverage_pca_prep) |> 
  rename(COVERAGEPC1 = "PC1", COVERAGEPC2 = "PC2", COVERAGEPC3 = "PC3")

coverage_pca_values_test <- bake(coverage_pca_prep, data_churn_test) |> 
  rename(COVERAGEPC1 = "PC1", COVERAGEPC2 = "PC2", COVERAGEPC3 = "PC3")

# SUBSCRIPTION

sub_pca_vars <- c('REVENUE','RECCHRGE','CHANGER','MONTHS','UNIQSUBS','ACTVSUBS')
#sub_pca_vars <- c('REVENUE','RECCHRGE')

sub_pca_data <- data_churn |>
  select(c('CUSTOMER','CHURN_TARGET',all_of(sub_pca_vars)))

sub_pca_rec <- recipe(~., data = sub_pca_data) %>%
  update_role(CUSTOMER, CHURN_TARGET, new_role = "id") %>%
  step_normalize(all_predictors()) %>%
  step_pca(all_predictors())

sub_pca_prep <- prep(sub_pca_rec)

sub_pca_values <- juice(sub_pca_prep) |> 
  rename(SUBPC1 = "PC1", SUBPC2 = "PC2", SUBPC3 = "PC3", SUBPC4 = "PC4", SUBPC5 = "PC5")

sub_pca_values_test <- bake(sub_pca_prep, data_churn_test) |> 
  rename(SUBPC1 = "PC1", SUBPC2 = "PC2", SUBPC3 = "PC3", SUBPC4 = "PC4", SUBPC5 = "PC5")

# CUSTOMER SERVICE

service_pca_vars <- c('CUSTCARE','MAILORD','MAILRES','MAILFLAG','RETCALLS','RETACCPT','REFER','RETCALL')

service_pca_data <- data_churn |>
  select(c('CUSTOMER','CHURN_TARGET',all_of(service_pca_vars)))

service_pca_rec <- recipe(~., data = service_pca_data) %>%
  update_role(CUSTOMER, CHURN_TARGET, new_role = "id") %>%
  step_normalize(all_predictors()) %>%
  step_pca(all_predictors())

service_pca_prep <- prep(service_pca_rec)

service_pca_values <- juice(service_pca_prep) |> 
  rename(SERVICEPC1 = "PC1", SERVICEPC2 = "PC2", SERVICEPC3 = "PC3", SERVICEPC4 = "PC4", SERVICEPC5 = "PC5")

service_pca_values_test <- bake(service_pca_prep, data_churn_test) |> 
  rename(SERVICEPC1 = "PC1", SERVICEPC2 = "PC2", SERVICEPC3 = "PC3", SERVICEPC4 = "PC4", SERVICEPC5 = "PC5")

# EQUIPMENT

eq_pca_vars <- c('PHONES','MODELS','EQPDAYS','REFURB','SETPRC')

eq_pca_data <- data_churn |>
  select(c('CUSTOMER','CHURN_TARGET',all_of(eq_pca_vars)))

eq_pca_rec <- recipe(~., data = eq_pca_data) %>%
  update_role(CUSTOMER, CHURN_TARGET, new_role = "id") %>%
  step_normalize(all_predictors()) %>%
  step_pca(all_predictors())

eq_pca_prep <- prep(eq_pca_rec)

eq_pca_values <- juice(eq_pca_prep) |> 
  rename(EQPC1 = "PC1", EQPC2 = "PC2", EQPC3 = "PC3", EQPC4 = "PC4", EQPC5 = "PC5")

eq_pca_values_test <- bake(eq_pca_prep, data_churn_test) |> 
  rename(EQPC1 = "PC1", EQPC2 = "PC2", EQPC3 = "PC3", EQPC4 = "PC4", EQPC5 = "PC5")

# PERSONAL

dem_pca_vars <- c('AGE1','AGE2','PCOWN','CREDITCD','INCOME')

dem_pca_data <- data_churn |>
  select(c('CUSTOMER','CHURN_TARGET',all_of(dem_pca_vars)))

dem_pca_rec <- recipe(~., data = dem_pca_data) %>%
  update_role(CUSTOMER, CHURN_TARGET, new_role = "id") %>%
  step_normalize(all_predictors()) %>%
  step_pca(all_predictors())

dem_pca_prep <- prep(dem_pca_rec)

dem_pca_values <- juice(dem_pca_prep) |> 
  rename(DEMPC1 = "PC1", DEMPC2 = "PC2", DEMPC3 = "PC3", DEMPC4 = "PC4", DEMPC5 = "PC5")

dem_pca_values_test <- bake(dem_pca_prep, data_churn_test) |> 
  rename(DEMPC1 = "PC1", DEMPC2 = "PC2", DEMPC3 = "PC3", DEMPC4 = "PC4", DEMPC5 = "PC5")

# ADD TO DATASETS

data_churn <- data_churn |> 
  left_join(usage_pca_values, by = 'CUSTOMER') |> 
  left_join(coverage_pca_values, by = 'CUSTOMER') |> 
  left_join(sub_pca_values, by = 'CUSTOMER') |> 
  left_join(service_pca_values, by = 'CUSTOMER') |> 
  left_join(eq_pca_values, by = 'CUSTOMER') |> 
  left_join(dem_pca_values, by = 'CUSTOMER')

data_churn_test <- data_churn_test |> 
  left_join(usage_pca_values_test, by = 'CUSTOMER') |> 
  left_join(coverage_pca_values_test, by = 'CUSTOMER') |> 
  left_join(sub_pca_values_test, by = 'CUSTOMER') |> 
  left_join(service_pca_values_test, by = 'CUSTOMER') |> 
  left_join(eq_pca_values_test, by = 'CUSTOMER') |> 
  left_join(dem_pca_values_test, by = 'CUSTOMER')

data_churn <- data_churn |> 
  mutate(USAGEPC1_high = case_when(USAGEPC1 >= 1 ~ 1, TRUE ~ 0)) |> 
  mutate(SERVICEPC2_mod = case_when(SERVICEPC2 >= -0.5 & SERVICEPC2 <= 1 ~ 1, TRUE ~ 0)) |> 
  mutate(EQPC4_low = case_when(EQPC4 <= 0.1 ~ 1, TRUE ~ 0)) |> 
  mutate(DEMPC5_low = case_when(DEMPC5 <= -0.5 ~ 1, TRUE ~ 0)) |> 
  mutate(EQPC2_low = case_when(EQPC2 <= -2 ~ 1, TRUE ~ 0)) |> 
  mutate(EQPC1_high = case_when(EQPC1 >= 0 ~ 1, TRUE ~ 0))



data_churn_test <- data_churn_test |> 
  mutate(USAGEPC1_high = case_when(USAGEPC1 >= 1 ~ 1, TRUE ~ 0)) |> 
  mutate(SERVICEPC2_mod = case_when(SERVICEPC2 >= -0.5 & SERVICEPC2 <= 1 ~ 1, TRUE ~ 0)) |> 
  mutate(EQPC4_low = case_when(EQPC4 <= 0.1 ~ 1, TRUE ~ 0)) |> 
  mutate(DEMPC5_low = case_when(DEMPC5 <= -0.5 ~ 1, TRUE ~ 0)) |> 
  mutate(EQPC2_low = case_when(EQPC2 <= -2 ~ 1, TRUE ~ 0)) |> 
  mutate(EQPC1_high = case_when(EQPC1 >= 0 ~ 1, TRUE ~ 0))

# summary(data_churn_test)
# data_churn_test |> 
#   ggplot(aes(x = DROPVCE, y = BLCKVCE, color = CHURN_TARGET)) +
#   geom_point()
```

## model variables

```{r}
model_vars <- c('CUSTOMER','CHURN_TARGET','MOU','RECCHRGE','CHANGEM_mod','DIRECTAS','CHANGER','CUSTCARE','MOUREC','INCALLS','PHONES','MODELS','SETPRC','CHILDREN','CREDITDEGY','PRIZMTWN','REFURB','SETPRCM','MARRY_LABEL','MAILRES','TRAVEL','PCOWN','RETCALL','MCYCLE', 'REFERRAL', 'CREDIT_ADJUSTMENT')#,'CSA_region_SEW','CSA_region_NS')

model_vars <- c('CUSTOMER','CHURN_TARGET','MOU','RECCHRGE','CHANGEM_mod','CHILDREN','CREDITDEGY','PRIZMTWN','REFURB','SETPRCM','MARRY_LABEL','MAILRES','TRAVEL','PCOWN','RETCALL','MCYCLE', 'REFERRAL', 'CREDIT_ADJUSTMENT','CSA_region_SEW','CSA_region_NS')

model_vars <- c('CUSTOMER','CHURN_TARGET','CHANGEM_mod','MOU_low','USAGEPC2','SUBPC5','SERVICEPC1','EQPC1','EQPC4_low','DEMPC5','MODELS','CHILDREN','CREDITDEGY','PRIZMTWN','REFURB','MARRYUN','MAILRES','TRAVEL','RETCALL','MCYCLE', 'REFERRAL', 'CREDIT_ADJUSTMENT','CSA_region_SEW','CSA_region_NS')

model_vars <- c('CUSTOMER','CHURN_TARGET','MOU','CHILDREN','CREDITDEGY','PRIZMTWN','TRAVEL','CSA_region_SEW','CSA_region_NS','CHANGEM_mod','ROAM','OVERAGE','DIRECTAS','CALLFWDV','REVENUE','RECCHRGE','RETCALL','MAILRES','REFER','PHONES','MODELS','SETPRC','REFURB','EQPDAYS','CHILDREN','TRAVEL','INCOME','PCOWN','CREDITCD','AGE1','AGE2')

model_vars <- c('CUSTOMER','CHURN_TARGET','MOU_low','CHILDREN','CREDITDEGY','PRIZMTWN','TRAVEL','CSA_region_SEW','CSA_region_NS','CHANGEM','USAGEPC2','RETCALL')

#model_vars <- c('PHONES','MODELS','SETPRC','EQPDAYS','REFURB')

working_model_data <- data_churn |>
  select(all_of(model_vars)) |> 
  select(!CUSTOMER)
```

## model for correlation & collinearity

```{r}
pairs_data <- working_model_data |> select(!c('CHURN_TARGET'))

res <- pairs_data |> cor()

corrplot(res, type = "upper", order = "hclust", tl.col = "black", tl.srt = 45)

model <- glm(CHURN_TARGET ~.,family=binomial(link='logit'),data=working_model_data)

vif(model)
```

## Regularized Logistic Regression Model

```{r}
set.seed(10)

train <- data_churn |>
  select(all_of(model_vars))

train <- upSample(x = train, y = train$CHURN_TARGET) |>
  select(!Class)

train$X <- sample(rep(c(0, 1, 2, 3), nrow(train) / 4))

train <- train |>
  distinct(CUSTOMER, X, .keep_all=TRUE) |>
  select(!c(X,CUSTOMER))

data_train_x <- model.matrix(CHURN_TARGET~., train)[,-1]
data_train_y <- as.integer(train$CHURN_TARGET)

cv.lasso <- cv.glmnet(data_train_x, data_train_y, alpha = 1, family = "binomial")

model1 <- glmnet(data_train_x, data_train_y, alpha = 1, family = "binomial",
                lambda = cv.lasso$lambda.min)

coef(model1)

data_test <- data_churn_test |>
  select(all_of(model_vars)) |> 
  select(!CUSTOMER)

x.test <- model.matrix(CHURN_TARGET ~., data_test)[,-1]
probabilities <- model1 %>% predict(newx = x.test,type="response")
predicted.classes <- ifelse(probabilities > 0.5, 1, 0)

observed.classes <- data_test$CHURN_TARGET
mean(predicted.classes == observed.classes)

pred_results <- data_test %>%
  select(CHURN_TARGET) %>%
  bind_cols(predicted.classes) |> 
  mutate(predicted = as.factor(s0))

conf_mat(pred_results,
         truth = CHURN_TARGET,
         estimate = predicted)

custom_metrics <- metric_set(accuracy, sens, yardstick::spec, yardstick::precision, yardstick::recall, f_meas, kap, mcc)

custom_metrics(pred_results,
               truth = CHURN_TARGET,
               estimate = predicted)
```

## explore PCs and variables

```{r, include = FALSE}
# usage_churn_data <- data_churn |>
#   select(c('CUSTOMER','CHURN_TARGET',all_of(usage_pca_vars)))
# 
# usage_pcas <- data_churn |>
#   select(c('CHURN_TARGET','USAGEPC1','USAGEPC2','USAGEPC3','USAGEPC4','USAGEPC5'))
# 
# usage_churn_data_test <- data_churn_test |>
#   select(c('CUSTOMER','CHURN_TARGET',all_of(usage_pca_vars)))
# 
# usage_pcas_test <- data_churn_test |>
#   select(c('CHURN_TARGET','USAGEPC1','USAGEPC2','USAGEPC3','USAGEPC4','USAGEPC5'))
# 
# 
# usage_tidied_pca <- tidy(usage_pca_prep, 2)
# 
# usage_tidied_pca %>%
#   filter(component %in% paste0("PC", 1:5)) %>%
#   mutate(component = fct_inorder(component)) %>%
#   ggplot(aes(value, terms, fill = terms)) +
#   geom_col(show.legend = FALSE) +
#   facet_wrap(~component, nrow = 1) +
#   labs(y = NULL)
# 
# usage_tidied_pca %>%
#   filter(component %in% paste0("PC", 1:5)) %>%
#   group_by(component) %>%
#   ungroup() %>%
#   mutate(terms = reorder_within(terms, abs(value), component)) %>%
#   ggplot(aes(abs(value), terms, fill = value > 0)) +
#   geom_col() +
#   facet_wrap(~component, scales = "free_y") +
#   scale_y_reordered() +
#   labs(x = "Absolute value of contribution",y = NULL, fill = "Positive?")
# 
# for(i in colnames(usage_pcas)) {
#   dist_df <- usage_pcas |>
#     select(c('CHURN_TARGET',all_of(i))) |>
#     na.omit()
#   print(i)
#   plot <- dist_df |>
#     ggplot(aes_string(x=names(dist_df)[2],color='CHURN_TARGET')) +
#     geom_density()
#   print(plot)
#   print('training')
# 
#   formula <- as.formula(paste(i, " ~ CHURN_TARGET"))
# 
#   if (i != 'CHURN_TARGET') {
#     ttest <- t.test(formula, dist_df)
#     print(ttest)
#   }
# 
#   dist_df <- usage_pcas_test |>
#     select(c('CHURN_TARGET',all_of(i))) |>
#     na.omit()
#   print(i)
#   plot <- dist_df |>
#     ggplot(aes_string(x=names(dist_df)[2],color='CHURN_TARGET')) +
#     geom_density()
#   print(plot)
#   print('training')
# 
#   formula <- as.formula(paste(i, " ~ CHURN_TARGET"))
# 
#   if (i != 'CHURN_TARGET') {
#     ttest <- t.test(formula, dist_df)
#     print(ttest)
#   }
# }
# 
# for(i in colnames(usage_churn_data)) {
#   dist_df <- usage_churn_data |>
#     select(c('CHURN_TARGET',all_of(i))) |>
#     na.omit()
#   print(i)
#   plot <- dist_df |>
#     ggplot(aes_string(x=names(dist_df)[2],color='CHURN_TARGET')) +
#     geom_density()
#   print(plot)
#   print('training')
# 
#   formula <- as.formula(paste(i, " ~ CHURN_TARGET"))
# 
#   if (i != 'CHURN_TARGET') {
#     ttest <- t.test(formula, dist_df)
#     print(ttest)
#   }
# 
#   dist_df <- usage_churn_data_test |>
#     select(c('CHURN_TARGET',all_of(i))) |>
#     na.omit()
#   print(i)
#   plot <- dist_df |>
#     ggplot(aes_string(x=names(dist_df)[2],color='CHURN_TARGET')) +
#     geom_density()
#   print(plot)
#   print('training')
# 
#   formula <- as.formula(paste(i, " ~ CHURN_TARGET"))
# 
#   if (i != 'CHURN_TARGET') {
#     ttest <- t.test(formula, dist_df)
#     print(ttest)
#   }
# }

# 
# 
# 
# 
# 
# 
# 
# 
# coverage_churn_data <- data_churn |> 
#   select(c('CUSTOMER','CHURN_TARGET',all_of(coverage_pca_vars)))
# 
# coverage_pcas <- data_churn |> 
#   select(c('CHURN_TARGET','COVERAGEPC1','COVERAGEPC2','COVERAGEPC3'))
# 
# coverage_churn_data_test <- data_churn_test |> 
#   select(c('CUSTOMER','CHURN_TARGET',all_of(coverage_pca_vars)))
# 
# coverage_pcas_test <- data_churn_test |> 
#   select(c('CHURN_TARGET','COVERAGEPC1','COVERAGEPC2','COVERAGEPC3'))
# 
# 
# coverage_tidied_pca <- tidy(coverage_pca_prep, 2)
# 
# coverage_tidied_pca %>%
#   filter(component %in% paste0("PC", 1:5)) %>%
#   mutate(component = fct_inorder(component)) %>%
#   ggplot(aes(value, terms, fill = terms)) +
#   geom_col(show.legend = FALSE) +
#   facet_wrap(~component, nrow = 1) +
#   labs(y = NULL)
# 
# coverage_tidied_pca %>%
#   filter(component %in% paste0("PC", 1:5)) %>%
#   group_by(component) %>%
#   ungroup() %>%
#   mutate(terms = reorder_within(terms, abs(value), component)) %>%
#   ggplot(aes(abs(value), terms, fill = value > 0)) +
#   geom_col() +
#   facet_wrap(~component, scales = "free_y") +
#   scale_y_reordered() +
#   labs(x = "Absolute value of contribution",y = NULL, fill = "Positive?")
# 
# for(i in colnames(coverage_pcas)) {
#   dist_df <- coverage_pcas |>
#     select(c('CHURN_TARGET',all_of(i))) |>
#     na.omit()
#   print(i)
#   plot <- dist_df |>
#     ggplot(aes_string(x=names(dist_df)[2],color='CHURN_TARGET')) +
#     geom_density()
#   print(plot)
#   print('training')
#   
#   formula <- as.formula(paste(i, " ~ CHURN_TARGET"))
#   
#   if (i != 'CHURN_TARGET') {
#     ttest <- t.test(formula, dist_df)
#     print(ttest)
#   }
#   
#   dist_df <- coverage_pcas_test |>
#     select(c('CHURN_TARGET',all_of(i))) |>
#     na.omit()
#   print(i)
#   plot <- dist_df |>
#     ggplot(aes_string(x=names(dist_df)[2],color='CHURN_TARGET')) +
#     geom_density()
#   print(plot)
#   print('training')
#   
#   formula <- as.formula(paste(i, " ~ CHURN_TARGET"))
#   
#   if (i != 'CHURN_TARGET') {
#     ttest <- t.test(formula, dist_df)
#     print(ttest)
#   }
# }
# 
# for(i in colnames(coverage_churn_data)) {
#   dist_df <- coverage_churn_data |>
#     select(c('CHURN_TARGET',all_of(i))) |>
#     na.omit()
#   print(i)
#   plot <- dist_df |>
#     ggplot(aes_string(x=names(dist_df)[2],color='CHURN_TARGET')) +
#     geom_density()
#   print(plot)
#   print('training')
#   
#   formula <- as.formula(paste(i, " ~ CHURN_TARGET"))
#   
#   if (i != 'CHURN_TARGET') {
#     ttest <- t.test(formula, dist_df)
#     print(ttest)
#   }
#   
#   dist_df <- coverage_churn_data_test |>
#     select(c('CHURN_TARGET',all_of(i))) |>
#     na.omit()
#   print(i)
#   plot <- dist_df |>
#     ggplot(aes_string(x=names(dist_df)[2],color='CHURN_TARGET')) +
#     geom_density()
#   print(plot)
#   print('training')
#   
#   formula <- as.formula(paste(i, " ~ CHURN_TARGET"))
#   
#   if (i != 'CHURN_TARGET') {
#     ttest <- t.test(formula, dist_df)
#     print(ttest)
#   }
# }
# 
# 
# 
# 
# 
# 
# 
# sub_churn_data <- data_churn |> 
#   select(c('CUSTOMER','CHURN_TARGET',all_of(sub_pca_vars)))
# 
# sub_pcas <- data_churn |> 
#   select(c('CHURN_TARGET','SUBPC1','SUBPC2','SUBPC3','SUBPC4','SUBPC5'))
# 
# sub_churn_data_test <- data_churn_test |> 
#   select(c('CUSTOMER','CHURN_TARGET',all_of(sub_pca_vars)))
# 
# sub_pcas_test <- data_churn_test |> 
#   select(c('CHURN_TARGET','SUBPC1','SUBPC2','SUBPC3','SUBPC4','SUBPC5'))
# 
# 
# sub_tidied_pca <- tidy(sub_pca_prep, 2)
# 
# sub_tidied_pca %>%
#   filter(component %in% paste0("PC", 1:5)) %>%
#   mutate(component = fct_inorder(component)) %>%
#   ggplot(aes(value, terms, fill = terms)) +
#   geom_col(show.legend = FALSE) +
#   facet_wrap(~component, nrow = 1) +
#   labs(y = NULL)
# 
# sub_tidied_pca %>%
#   filter(component %in% paste0("PC", 1:5)) %>%
#   group_by(component) %>%
#   ungroup() %>%
#   mutate(terms = reorder_within(terms, abs(value), component)) %>%
#   ggplot(aes(abs(value), terms, fill = value > 0)) +
#   geom_col() +
#   facet_wrap(~component, scales = "free_y") +
#   scale_y_reordered() +
#   labs(x = "Absolute value of contribution",y = NULL, fill = "Positive?")
# 
# for(i in colnames(sub_pcas)) {
#   dist_df <- sub_pcas |>
#     select(c('CHURN_TARGET',all_of(i))) |>
#     na.omit()
#   print(i)
#   plot <- dist_df |>
#     ggplot(aes_string(x=names(dist_df)[2],color='CHURN_TARGET')) +
#     geom_density()
#   print(plot)
#   print('training')
#   
#   formula <- as.formula(paste(i, " ~ CHURN_TARGET"))
#   
#   if (i != 'CHURN_TARGET') {
#     ttest <- t.test(formula, dist_df)
#     print(ttest)
#   }
#   
#   dist_df <- sub_pcas_test |>
#     select(c('CHURN_TARGET',all_of(i))) |>
#     na.omit()
#   print(i)
#   plot <- dist_df |>
#     ggplot(aes_string(x=names(dist_df)[2],color='CHURN_TARGET')) +
#     geom_density()
#   print(plot)
#   print('training')
#   
#   formula <- as.formula(paste(i, " ~ CHURN_TARGET"))
#   
#   if (i != 'CHURN_TARGET') {
#     ttest <- t.test(formula, dist_df)
#     print(ttest)
#   }
# }
# 
# for(i in colnames(sub_churn_data)) {
#   dist_df <- sub_churn_data |>
#     select(c('CHURN_TARGET',all_of(i))) |>
#     na.omit()
#   print(i)
#   plot <- dist_df |>
#     ggplot(aes_string(x=names(dist_df)[2],color='CHURN_TARGET')) +
#     geom_density()
#   print(plot)
#   print('training')
#   
#   formula <- as.formula(paste(i, " ~ CHURN_TARGET"))
#   
#   if (i != 'CHURN_TARGET') {
#     ttest <- t.test(formula, dist_df)
#     print(ttest)
#   }
#   
#   dist_df <- sub_churn_data_test |>
#     select(c('CHURN_TARGET',all_of(i))) |>
#     na.omit()
#   print(i)
#   plot <- dist_df |>
#     ggplot(aes_string(x=names(dist_df)[2],color='CHURN_TARGET')) +
#     geom_density()
#   print(plot)
#   print('training')
#   
#   formula <- as.formula(paste(i, " ~ CHURN_TARGET"))
#   
#   if (i != 'CHURN_TARGET') {
#     ttest <- t.test(formula, dist_df)
#     print(ttest)
#   }
# }
# 
# 
# 
# 
# 
# 
# service_churn_data <- data_churn |> 
#   select(c('CUSTOMER','CHURN_TARGET',all_of(service_pca_vars)))
# 
# service_pcas <- data_churn |> 
#   select(c('CHURN_TARGET','SERVICEPC1','SERVICEPC2','SERVICEPC3','SERVICEPC4','SERVICEPC5'))
# 
# service_churn_data_test <- data_churn_test |> 
#   select(c('CUSTOMER','CHURN_TARGET',all_of(service_pca_vars)))
# 
# service_pcas_test <- data_churn_test |> 
#   select(c('CHURN_TARGET','SERVICEPC1','SERVICEPC2','SERVICEPC3','SERVICEPC4','SERVICEPC5'))
# 
# 
# service_tidied_pca <- tidy(service_pca_prep, 2)
# 
# service_tidied_pca %>%
#   filter(component %in% paste0("PC", 1:5)) %>%
#   mutate(component = fct_inorder(component)) %>%
#   ggplot(aes(value, terms, fill = terms)) +
#   geom_col(show.legend = FALSE) +
#   facet_wrap(~component, nrow = 1) +
#   labs(y = NULL)
# 
# service_tidied_pca %>%
#   filter(component %in% paste0("PC", 1:5)) %>%
#   group_by(component) %>%
#   ungroup() %>%
#   mutate(terms = reorder_within(terms, abs(value), component)) %>%
#   ggplot(aes(abs(value), terms, fill = value > 0)) +
#   geom_col() +
#   facet_wrap(~component, scales = "free_y") +
#   scale_y_reordered() +
#   labs(x = "Absolute value of contribution",y = NULL, fill = "Positive?")
# 
# for(i in colnames(service_pcas)) {
#   dist_df <- service_pcas |>
#     select(c('CHURN_TARGET',all_of(i))) |>
#     na.omit()
#   print(i)
#   plot <- dist_df |>
#     ggplot(aes_string(x=names(dist_df)[2],color='CHURN_TARGET')) +
#     geom_density()
#   print(plot)
#   print('training')
#   
#   formula <- as.formula(paste(i, " ~ CHURN_TARGET"))
#   
#   if (i != 'CHURN_TARGET') {
#     ttest <- t.test(formula, dist_df)
#     print(ttest)
#   }
#   
#   dist_df <- service_pcas_test |>
#     select(c('CHURN_TARGET',all_of(i))) |>
#     na.omit()
#   print(i)
#   plot <- dist_df |>
#     ggplot(aes_string(x=names(dist_df)[2],color='CHURN_TARGET')) +
#     geom_density()
#   print(plot)
#   print('training')
#   
#   formula <- as.formula(paste(i, " ~ CHURN_TARGET"))
#   
#   if (i != 'CHURN_TARGET') {
#     ttest <- t.test(formula, dist_df)
#     print(ttest)
#   }
# }
# 
# for(i in colnames(service_churn_data)) {
#   dist_df <- service_churn_data |>
#     select(c('CHURN_TARGET',all_of(i))) |>
#     na.omit()
#   print(i)
#   plot <- dist_df |>
#     ggplot(aes_string(x=names(dist_df)[2],color='CHURN_TARGET')) +
#     geom_density()
#   print(plot)
#   print('training')
#   
#   formula <- as.formula(paste(i, " ~ CHURN_TARGET"))
#   
#   if (i != 'CHURN_TARGET') {
#     ttest <- t.test(formula, dist_df)
#     print(ttest)
#   }
#   
#   dist_df <- service_churn_data_test |>
#     select(c('CHURN_TARGET',all_of(i))) |>
#     na.omit()
#   print(i)
#   plot <- dist_df |>
#     ggplot(aes_string(x=names(dist_df)[2],color='CHURN_TARGET')) +
#     geom_density()
#   print(plot)
#   print('training')
#   
#   formula <- as.formula(paste(i, " ~ CHURN_TARGET"))
#   
#   if (i != 'CHURN_TARGET') {
#     ttest <- t.test(formula, dist_df)
#     print(ttest)
#   }
# }
# 
# 
# 
# 
# 
# 
# 

# eq_churn_data <- data_churn |>
#   select(c('CUSTOMER','CHURN_TARGET',all_of(eq_pca_vars)))
# 
# eq_pcas <- data_churn |>
#   select(c('CHURN_TARGET','EQPC1','EQPC2','EQPC3','EQPC4','EQPC5'))
# 
# eq_churn_data_test <- data_churn_test |>
#   select(c('CUSTOMER','CHURN_TARGET',all_of(eq_pca_vars)))
# 
# eq_pcas_test <- data_churn_test |>
#   select(c('CHURN_TARGET','EQPC1','EQPC2','EQPC3','EQPC4','EQPC5'))
# 
# 
# eq_tidied_pca <- tidy(eq_pca_prep, 2)
# 
# eq_tidied_pca %>%
#   filter(component %in% paste0("PC", 1:5)) %>%
#   mutate(component = fct_inorder(component)) %>%
#   ggplot(aes(value, terms, fill = terms)) +
#   geom_col(show.legend = FALSE) +
#   facet_wrap(~component, nrow = 1) +
#   labs(y = NULL)
# 
# eq_tidied_pca %>%
#   filter(component %in% paste0("PC", 1:5)) %>%
#   group_by(component) %>%
#   ungroup() %>%
#   mutate(terms = reorder_within(terms, abs(value), component)) %>%
#   ggplot(aes(abs(value), terms, fill = value > 0)) +
#   geom_col() +
#   facet_wrap(~component, scales = "free_y") +
#   scale_y_reordered() +
#   labs(x = "Absolute value of contribution",y = NULL, fill = "Positive?")
# 
# for(i in colnames(eq_pcas)) {
#   dist_df <- eq_pcas |>
#     select(c('CHURN_TARGET',all_of(i))) |>
#     na.omit()
#   print(i)
#   plot <- dist_df |>
#     ggplot(aes_string(x=names(dist_df)[2],color='CHURN_TARGET')) +
#     geom_density()
#   print(plot)
#   print('training')
# 
#   formula <- as.formula(paste(i, " ~ CHURN_TARGET"))
# 
#   if (i != 'CHURN_TARGET') {
#     ttest <- t.test(formula, dist_df)
#     print(ttest)
#   }
# 
#   dist_df <- eq_pcas_test |>
#     select(c('CHURN_TARGET',all_of(i))) |>
#     na.omit()
#   print(i)
#   plot <- dist_df |>
#     ggplot(aes_string(x=names(dist_df)[2],color='CHURN_TARGET')) +
#     geom_density()
#   print(plot)
#   print('training')
# 
#   formula <- as.formula(paste(i, " ~ CHURN_TARGET"))
# 
#   if (i != 'CHURN_TARGET') {
#     ttest <- t.test(formula, dist_df)
#     print(ttest)
#   }
# }
# 
# for(i in colnames(eq_churn_data)) {
#   dist_df <- eq_churn_data |>
#     select(c('CHURN_TARGET',all_of(i))) |>
#     na.omit()
#   print(i)
#   plot <- dist_df |>
#     ggplot(aes_string(x=names(dist_df)[2],color='CHURN_TARGET')) +
#     geom_density()
#   print(plot)
#   print('training')
# 
#   formula <- as.formula(paste(i, " ~ CHURN_TARGET"))
# 
#   if (i != 'CHURN_TARGET') {
#     ttest <- t.test(formula, dist_df)
#     print(ttest)
#   }
# 
#   dist_df <- eq_churn_data_test |>
#     select(c('CHURN_TARGET',all_of(i))) |>
#     na.omit()
#   print(i)
#   plot <- dist_df |>
#     ggplot(aes_string(x=names(dist_df)[2],color='CHURN_TARGET')) +
#     geom_density()
#   print(plot)
#   print('training')
# 
#   formula <- as.formula(paste(i, " ~ CHURN_TARGET"))
# 
#   if (i != 'CHURN_TARGET') {
#     ttest <- t.test(formula, dist_df)
#     print(ttest)
#   }
# }
# 
# 
# 
# 
# 
#
```


```{r}
dem_churn_data <- data_churn |>
  select(c('CUSTOMER','CHURN_TARGET',all_of(dem_pca_vars)))

dem_pcas <- data_churn |>
  select(c('CHURN_TARGET','DEMPC1','DEMPC2','DEMPC3','DEMPC4','DEMPC5'))

dem_churn_data_test <- data_churn_test |>
  select(c('CUSTOMER','CHURN_TARGET',all_of(dem_pca_vars)))

dem_pcas_test <- data_churn_test |>
  select(c('CHURN_TARGET','DEMPC1','DEMPC2','DEMPC3','DEMPC4','DEMPC5'))


dem_tidied_pca <- tidy(dem_pca_prep, 2)

dem_tidied_pca %>%
  filter(component %in% paste0("PC", 1:5)) %>%
  mutate(component = fct_inorder(component)) %>%
  ggplot(aes(value, terms, fill = terms)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~component, nrow = 1) +
  labs(y = NULL)

dem_tidied_pca %>%
  filter(component %in% paste0("PC", 1:5)) %>%
  group_by(component) %>%
  ungroup() %>%
  mutate(terms = reorder_within(terms, abs(value), component)) %>%
  ggplot(aes(abs(value), terms, fill = value > 0)) +
  geom_col() +
  facet_wrap(~component, scales = "free_y") +
  scale_y_reordered() +
  labs(x = "Absolute value of contribution",y = NULL, fill = "Positive?")

for(i in colnames(dem_pcas)) {
  dist_df <- dem_pcas |>
    select(c('CHURN_TARGET',all_of(i))) |>
    na.omit()
  print(i)
  plot <- dist_df |>
    ggplot(aes_string(x=names(dist_df)[2],color='CHURN_TARGET')) +
    geom_density()
  print(plot)
  print('training')

  formula <- as.formula(paste(i, " ~ CHURN_TARGET"))

  if (i != 'CHURN_TARGET') {
    ttest <- t.test(formula, dist_df)
    print(ttest)
  }

  dist_df <- dem_pcas_test |>
    select(c('CHURN_TARGET',all_of(i))) |>
    na.omit()
  print(i)
  plot <- dist_df |>
    ggplot(aes_string(x=names(dist_df)[2],color='CHURN_TARGET')) +
    geom_density()
  print(plot)
  print('training')

  formula <- as.formula(paste(i, " ~ CHURN_TARGET"))

  if (i != 'CHURN_TARGET') {
    ttest <- t.test(formula, dist_df)
    print(ttest)
  }
}

for(i in colnames(dem_churn_data)) {
  dist_df <- dem_churn_data |>
    select(c('CHURN_TARGET',all_of(i))) |>
    na.omit()
  print(i)
  plot <- dist_df |>
    ggplot(aes_string(x=names(dist_df)[2],color='CHURN_TARGET')) +
    geom_density()
  print(plot)
  print('training')

  formula <- as.formula(paste(i, " ~ CHURN_TARGET"))

  if (i != 'CHURN_TARGET') {
    ttest <- t.test(formula, dist_df)
    print(ttest)
  }

  dist_df <- dem_churn_data_test |>
    select(c('CHURN_TARGET',all_of(i))) |>
    na.omit()
  print(i)
  plot <- dist_df |>
    ggplot(aes_string(x=names(dist_df)[2],color='CHURN_TARGET')) +
    geom_density()
  print(plot)
  print('training')

  formula <- as.formula(paste(i, " ~ CHURN_TARGET"))

  if (i != 'CHURN_TARGET') {
    ttest <- t.test(formula, dist_df)
    print(ttest)
  }
}
```

## model variables

```{r}
model_vars <- c('CUSTOMER','CHURN_TARGET','MOU','RECCHRGE','CHANGEM_mod','DIRECTAS','CHANGER','CUSTCARE','MOUREC','INCALLS','PHONES','MODELS','SETPRC','CHILDREN','CREDITDEGY','PRIZMTWN','REFURB','SETPRCM','MARRY_LABEL','MAILRES','TRAVEL','PCOWN','RETCALL','MCYCLE', 'REFERRAL', 'CREDIT_ADJUSTMENT')#,'CSA_region_SEW','CSA_region_NS')

model_vars <- c('CUSTOMER','CHURN_TARGET','MOU','RECCHRGE','CHANGEM_mod','CHILDREN','CREDITDEGY','PRIZMTWN','REFURB','SETPRCM','MARRY_LABEL','MAILRES','TRAVEL','PCOWN','RETCALL','MCYCLE', 'REFERRAL', 'CREDIT_ADJUSTMENT','CSA_region_SEW','CSA_region_NS')

model_vars <- c('CUSTOMER','CHURN_TARGET','CHANGEM_mod','MOU_low','USAGEPC2','SUBPC5','SERVICEPC1','EQPC1','EQPC4_low','DEMPC5','MODELS','CHILDREN','CREDITDEGY','PRIZMTWN','REFURB','MARRYUN','MAILRES','TRAVEL','RETCALL','MCYCLE', 'REFERRAL', 'CREDIT_ADJUSTMENT','CSA_region_SEW','CSA_region_NS')

model_vars <- c('CUSTOMER','CHURN_TARGET','CHANGEM_mod','MOU_low','MODELS','CHILDREN','CREDITDEGY','PRIZMTWN','REFURB','MARRYUN','MAILRES','TRAVEL','RETCALL', 'REFERRAL', 'CSA_region_SEW', 'CSA_region_NS')

working_model_data <- data_churn |>
  select(all_of(model_vars))
```

## variables by category

```{r}
#5
id_vars <- c('CUSTOMER','OCC','MARRY')
#1
target_vars <- c('CHURN')
#16
usage_vars <- c('MOU','OVERAGE','ROAM','CHANGEM','UNANSVCE','THREEWAY','MOUREC','OUTCALLS','INCALLS','PEAKVCE','OPEAKVCE','CALLFWDV','CALLWAIT','NEWCELLY','NEWCELLN')
#2
coverage_vars <- c('DROPVCE','DROPBLK','BLCKVCE')
#6
sub_vars <- c('REVENUE','RECCHRGE','CHANGER','MONTHS','UNIQSUBS','ACTVSUBS')
#9
service_vars <- c('DIRECTAS','CUSTCARE','MAILORD','MAILRES','MAILFLAG','RETCALLS','RETACCPT','REFER','RETCALL')
#7
eq_vars <- c('PHONES','MODELS','EQPDAYS','REFURB','WEBCAP','SETPRCM','SETPRC')
#4
dem_vars_label <- c('CREDIT_RATING','PRZM_NUM','OCC_LABEL','MARRY_LABEL')
#20
dem_vars_dich <- c('CREDITA', 'CREDITAA', 'CREDITB', 'CREDITC', 'CREDITDE', 'CREDITGY', 'CREDITZ', 'PRIZMRUR', 'PRIZMUB', 'PRIZMTWN', 'OCCPROF', 'OCCCLER', 'OCCCRFT', 'OCCSTUD', 'OCCHMKR', 'OCCRET', 'OCCSELF', 'MARRYUN', 'MARRYYES', 'MARRYNO')
#13
dem_vars_other <- c('AGE1','AGE2','CHILDREN','TRUCK','RV','OWNRENT','TRAVEL','PCOWN','CREDITCD','INCMISS','INCOME','MCYCLE','CREDITAD')
#2
location_vars <- c('CSA','CSA_main')
```

## PCA Data

```{r}
pca_data <- data_churn
#5
id_pca_vars <- c('CUSTOMER')
#1
target_pca_vars <- c('CHURN_TARGET')
#16
usage_pca_vars <- c('MOU','DIRECTAS','OVERAGE','ROAM','CHANGEM','DROPVCE','BLCKVCE','UNANSVCE','THREEWAY','MOUREC','OUTCALLS','INCALLS','PEAKVCE','OPEAKVCE','DROPBLK','CALLFWDV','CALLWAIT')
#2
coverage_pca_vars <- c('DROPVCE','DROPBLK','BLCKVCE')
#6
sub_pca_vars <- c('REVENUE','RECCHRGE','CHANGER','MONTHS','UNIQSUBS','ACTVSUBS')
#sub_pca_vars <- c('REVENUE','RECCHRGE')
#9
service_pca_vars <- c('CUSTCARE','MAILORD','MAILRES','MAILFLAG','RETCALLS','RETACCPT','REFER','RETCALL')
#7
eq_pca_vars <- c('PHONES','MODELS','EQPDAYS','REFURB','WEBCAP','SETPRCM','SETPRC','NEWCELLY','NEWCELLN')
#13
dem_pca_vars <- c('AGE1','AGE2','PCOWN','CREDITCD','INCOME','MCYCLE','CREDITAD')
```

## variables by type

```{r}
target_vars <- c('CUSTOMER','CHURN_TARGET','CHURN_INT')

numeric_vars <- c('REVENUE','MOU','RECCHRGE','DIRECTAS','OVERAGE','ROAM','CHANGEM','CHANGER','DROPVCE','BLCKVCE','UNANSVCE','CUSTCARE','THREEWAY','MOUREC','OUTCALLS','INCALLS','PEAKVCE','OPEAKVCE','DROPBLK','CALLFWDV','CALLWAIT','MONTHS','UNIQSUBS','ACTVSUBS','PHONES','MODELS','EQPDAYS','AGE1','AGE2','INCOME','SETPRC','USAGEPC1','USAGEPC2','USAGEPC3','USAGEPC4','COVERAGEPC1','COVERAGEPC2','COVERAGEPC3','SUBPC1','SUBPC2','SUBPC3','SUBPC4','SUBPC5','SERVICEPC1','SERVICEPC2','SERVICEPC3','SERVICEPC4','SERVICEPC5','EQPC1','EQPC2','EQPC3','EQPC4','EQPC5','DEMPC1','DEMPC2','DEMPC3','DEMPC4','DEMPC5')

categorical_vars <- c('INCOME_medhigh','CREDITDEGY', 'REFERRAL','CREDIT_ADJUSTMENT','CHILDREN','CREDIT_LABEL','PRIZM_LABEL','REFURB','WEBCAP','TRUCK','RV','OCC_LABEL','OWNRENT','INCMISS','SETPRCM','MARRY_LABEL','MAILORD','MAILRES','MAILFLAG','TRAVEL','PCOWN','CREDITCD','RETCALLS','RETACCPT','NEWCELLY','NEWCELLN','REFER','MCYCLE','CREDITAD','RETCALL', 'CSA_state', 'CSA_region','MOU_low','CHANGEM_mod', 'USAGEPC1_high','SERVICEPC2_mod','EQPC4_low','DEMPC5_low','EQPC2_low','EQPC1_high')

data_churn <- data_churn |> 
  mutate(CREDITDEGY = as.factor(CREDITDEGY), CHILDREN = as.factor(CHILDREN), CREDIT_LABEL = as.factor(CREDIT_LABEL), PRIZM_LABEL = as.factor(PRIZM_LABEL), REFURB = as.factor(REFURB), WEBCAP = as.factor(WEBCAP), TRUCK = as.factor(TRUCK), RV = as.factor(RV), OCC_LABEL = as.factor(OCC_LABEL), OWNRENT = as.factor(OWNRENT), INCMISS = as.factor(INCMISS), SETPRCM = as.factor(SETPRCM), MARRY_LABEL = as.factor(MARRY_LABEL), MAILORD = as.factor(MAILORD), MAILRES = as.factor(MAILRES), MAILFLAG = as.factor(MAILFLAG), TRAVEL = as.factor(TRAVEL), PCOWN = as.factor(PCOWN), CREDITCD = as.factor(CREDITCD), RETCALLS = as.factor(RETCALLS), RETACCPT = as.factor(RETACCPT), NEWCELLY = as.factor(NEWCELLY), NEWCELLN = as.factor(NEWCELLN), REFER = as.factor(REFER), MCYCLE = as.factor(MCYCLE), CREDITAD = as.factor(CREDITAD), RETCALL = as.factor(RETCALL), MOU_low = as.factor(MOU_low), USAGEPC1_high = as.factor(USAGEPC1_high), SERVICEPC2_mod = as.factor(SERVICEPC2_mod), EQPC4_low = as.factor(EQPC4_low), DEMPC5_low = as.factor(DEMPC5_low), CHANGEM_mod = as.factor(CHANGEM_mod), EQPC2_low = as.factor(EQPC2_low), EQPC1_high = as.factor(EQPC1_high))
```

## data distributions

```{r}
churn_data <- data_churn |>
  select(all_of(target_vars), all_of(numeric_vars), all_of(categorical_vars))
# 
# numeric_df <- churn_data |> 
#   select(c(CHURN_TARGET,all_of(numeric_vars)))
# 
# numeric_df_test <- data_churn_test |> 
#   select(c(CHURN_TARGET,all_of(numeric_vars)))
# 
# for(i in colnames(numeric_df)) {
#   dist_df <- numeric_df |>
#     select(c('CHURN_TARGET',all_of(i))) |>
#     na.omit()
#   print(i)
#   plot <- dist_df |>
#     ggplot(aes_string(x=names(dist_df)[2],color='CHURN_TARGET')) +
#     geom_density()
#   print(plot)
#   print('training')
#   
#   formula <- as.formula(paste(i, " ~ CHURN_TARGET"))
#   
#   if (i != 'CHURN_TARGET') {
#     ttest <- t.test(formula, dist_df)
#     print(ttest)
#   }
#   
#   dist_df <- numeric_df_test |>
#     select(c('CHURN_TARGET',all_of(i))) |>
#     na.omit()
#   plot <- dist_df |>
#     ggplot(aes_string(x=names(dist_df)[2],color='CHURN_TARGET')) +
#     geom_density()
#   print(plot)
#   print('testing')
#   
#   formula <- as.formula(paste(i, " ~ CHURN_TARGET"))
#   if (i != 'CHURN_TARGET') {
#     ttest <- t.test(formula, dist_df)
#     print(ttest)
#   }
# }
# 
categorical_df <- data_churn |>
  select(c(CHURN_INT,CHURN_TARGET,all_of(categorical_vars)))

categorical_df_test <- data_churn_test |>
  select(c(CHURN_INT,CHURN_TARGET,all_of(categorical_vars)))

for(i in colnames(categorical_df)) {
  dist_df <- categorical_df |>
    select(c('CHURN_INT','CHURN_TARGET',all_of(i))) |>
    na.omit()
  print(i)
  plot <- dist_df |>
    ggplot(aes_string(x=names(dist_df)[3],fill='CHURN_TARGET')) +
    facet_wrap('CHURN_TARGET', nrow = 2) +
    geom_bar()
  #print(plot)
  print('training')

  dist_df$X <- pull(dist_df[,which(colnames(dist_df)==i)])
  dist_grouped <- dist_df |>
    group_by(X) |>
    summarize(count = n(), churn = sum(CHURN_INT)) |>
    mutate(pct = churn / count)
  print(dist_grouped)

  dist_df <- categorical_df_test |>
    select(c('CHURN_INT','CHURN_TARGET',all_of(i))) |>
    na.omit()
  plot <- dist_df |>
    ggplot(aes_string(x=names(dist_df)[3],fill='CHURN_TARGET')) +
    facet_wrap('CHURN_TARGET', nrow = 2) +
    geom_bar()
  #print(plot)
  print('testing')

  dist_df$X <- pull(dist_df[,which(colnames(dist_df)==i)])
  dist_grouped <- dist_df |>
    group_by(X) |>
    summarize(count = n(), churn = sum(CHURN_INT)) |>
    mutate(pct = churn / count)
  print(dist_grouped)
}
```

## t tests

```{r}
# for(i in colnames(numeric_df)) {
#   dist_df <- numeric_df |>
#     select(c('CHURN',all_of(i))) |>
#     na.omit()
#   print(i)
#   formula <- as.formula(paste(i, " ~ CHURN"))
#   if (i != 'CHURN') {
#     ttest <- t.test(formula, dist_df)
#     print(ttest)
#   }
# }
# 
# t.test(REVENUE ~ CHURN, numeric_df)
```

## selecting model variables

```{r}
#model_vars <- c('CUSTOMER','CHURN_TARGET','MOU','MOUREC','OUTCALLS','DIRECTAS','CHANGEM','PEAKVCE','CALLWAIT','OPEAKVCE','REVENUE','RECCHRGE','DROPBLK','UNANSVCE','ACTVSUBS','MAILRES','RETCALLS','EQPDAYS','REFURB')

# model_vars <- c('CUSTOMER','CHURN_TARGET','MOU','OPEAKVCE','RETCALLS','EQPDAYS','REFURB','CSA_region','CSA_region_SEW','CSA_region_SW')
# 
# model_vars <- c('CUSTOMER','CHURN','MOU','OVERAGE','ROAM','CHANGEM','UNANSVCE','THREEWAY','MOUREC','OUTCALLS','INCALLS','PEAKVCE','OPEAKVCE','CALLFWDV','CALLWAIT','NEWCELLY','NEWCELLN','DROPVCE','DROPBLK','BLCKVCE','REVENUE','RECCHRGE','CHANGER','MONTHS','UNIQSUBS','ACTVSUBS','DIRECTAS','CUSTCARE','MAILORD','MAILRES','MAILFLAG','RETCALLS','RETACCPT','REFER','RETCALL','PHONES','MODELS','EQPDAYS','REFURB','WEBCAP','SETPRC','CREDIT_RATING','PRZM_NUM','OCC_LABEL','MARRYUN','MARRYYES','MARRYNO','AGE1','AGE2','CHILDREN','TRUCK','RV','OWNRENT','TRAVEL','PCOWN','CREDITCD','INCMISS','INCOME','MCYCLE','CREDITAD','CSA_main')
# 
# model_vars <- c('CUSTOMER','CHURN_TARGET','MOU','OVERAGE','ROAM','CHANGEM','UNANSVCE','THREEWAY','MOUREC','OUTCALLS','INCALLS','PEAKVCE','OPEAKVCE','CALLFWDV','CALLWAIT','NEWCELLY','NEWCELLN','DROPVCE','DROPBLK','BLCKVCE','REVENUE','RECCHRGE','CHANGER','MONTHS','UNIQSUBS','ACTVSUBS','DIRECTAS','CUSTCARE','MAILORD','MAILRES','MAILFLAG','RETCALLS','RETACCPT','REFER','RETCALL','PHONES','MODELS','EQPDAYS','REFURB','WEBCAP','SETPRC','CREDIT_LABEL','PRIZM_LABEL','MARRY_LABEL','AGE1','AGE2','CHILDREN','TRUCK','RV','TRAVEL','PCOWN','CREDITCD','INCOME','MCYCLE','CREDITAD')
```



