---
title: "Surdek Milestone 3"
author: "msurdek"
date: "7/17/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(width = 120)
```

```{r, message=FALSE, warnings=FALSE, results=FALSE}
library(dplyr)
library(readr)
library(randomForest)
library(caret)
library(scales)
```

```{r, message=FALSE, warnings=FALSE}
data <- read_csv("/home/msurdek/Documents/School/SNHU/DAT 650 Advanced Data Analytics/Assignments/Project//Data/Cell Data.csv") |> 
  filter(CALIBRAT == 1) |> 
  mutate(CHURNDEP = as.factor(CHURNDEP)) |> 
  rename(PRZM_NUM = 'Column 45') |> 
  mutate(CSA_main = substr(CSA,1,3)) |> 
  na.omit()

cell_data <- data |>
  select(!c(CUSTOMER,CHURN)) |> 
  select(!c(CREDITA, CREDITAA,  CREDITB, CREDITC, CREDITDE, CREDITGY, CREDITZ)) |> 
  select(!c(PRIZMRUR, PRIZMUB, PRIZMTWN)) |> 
  select(!c(OCCPROF, OCCCLER, OCCCRFT, OCCSTUD, OCCHMKR, OCCRET, OCCSELF, OCC)) |> 
  select(!c(MARRYUN, MARRYYES, MARRYNO, MARRY))
```

```{r}
cellDS <- new.env()

evalq({
  data <- na.omit(cell_data)
  nobs <- nrow(data)
  form <- formula(CHURNDEP ~ .)
  target <- all.vars(form)[1]
  vars <- names(data)
  set.seed(10)
  train <- sample(nobs, 1 * nobs)
  training <- data[train,]
}, cellDS)

cellRF <- new.env(parent = cellDS)

evalq({
  model <- randomForest(formula = form,
                        data = data[train, vars],
                        ntree = 100, mtry = 6,
                        importance = TRUE,
                        localImp =TRUE,
                        replace = FALSE)
}, cellRF)
```


model <- randomForest(formula = form,
                      data = data[train, vars],
                      ntree = 100, mtry = 6,
                      importance = TRUE,
                      localImp =TRUE,
                      replace = FALSE)



```{r}
str(cellRF$model)

cellRF$model

head(cellRF$model$predicted, 10)

head(cellRF$model$importance)

data.frame(round(cellRF$model$importance,4)) |> 
  arrange(desc(MeanDecreaseAccuracy))

data.frame(round(cellRF$model$importance,4)) |> 
  arrange(desc(MeanDecreaseGini))

head(cellRF$model$localImp)[,1:4]

cellRF$model$err.rate

round(head(cellRF$model$err.rate, 15), 4)

evalq({
  min.err <- min(data.frame(model$err.rate)["OOB"])
  min.err.idx <- which(data.frame(model$err.rate)["OOB"] == min.err)
}, cellRF)

cellRF$min.err

cellRF$min.err.idx

cellRF$model$err.rate[cellRF$min.err.idx]

votes <- data.frame(cellRF$model$votes) |> 
  cbind(data$CUSTOMER)|> 
  select(c('data$CUSTOMER','X1'))

colnames(votes) <- c('CUSTOMER_ID','CHURN_RISK')

votes |>  
  arrange(desc(CHURN_RISK))  |>
  mutate(CHURN_RISK = label_percent(accuracy=.01)(CHURN_RISK)) |> 
  head(10) |> 
  print(row.names = FALSE, quote = FALSE)

head(apply(cellRF$model$votes, 1, sum))

plot(cellRF$model)

confusionMatrix(cellRF$model$predicted, cellDS$training$CHURNDEP)

#cellDS$predval <- predict(cellRF$model, cellDS$validation)
#confusionMatrix(cellDS$predval, cellDS$validation$CHURNDEP)

```
